---
title: "Simplifying Stats Jargon"
author: "Your Name"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE,
                      fig.height = 8, fig.width = 12)
```

## Introduction

The following is a quick "helper" to aid in understanding stats jargon.

# Describing Groups - Calculating Descriptive Statistics

Useful ways to describe groups include measures of "center" eg. mean or median and measures of "spread" eg. variance, standard deviation or standard error.

* If you don't think your data is normal
    - Use standard deviation (describes variation in data). Don't us standard error (describes variation about mean).
    - Use median because not as effected by funky distributions
    

```{r}
library(tidyverse)
set.seed(123)
data <- tibble(x = runif(20),y=runif(20),a=rep(c("A","B"),each=10))

desc_table <- data %>%
  group_by(a) %>%
  summarise(
    n = n(),
    mean = mean(y),
    sd = sd(y),
    se = sd/sqrt(n)
  )

#Note that the se is used to compare means (variation about mean taking into account sample size, so if you have a huuuuuge sample the confidence in the mean shrinks).
#Also note that se standard deviation is the sqrt of variance which is is the squared sum of diff between points and the mean over n-1 (bessels correction). n subracts one because we know the sample mean and thus can calculate the "last" n.

#There are MANY ways to show the measures of center and variation. This is one way using SUMMARIZE FIRST.
desc_table %>%
  ggplot(aes(x = a,color=a)) +
  geom_errorbar(aes(ymin=mean-se,ymax=mean+se)) +
  geom_point(aes(y=mean),size=3, shape=21, fill="white")

#This is one way using SUMMARIZE IN GGPLOT2.
#TIP: use CMD + click on function to get the details..eg. for mean_se
data %>%
  ggplot(aes(x = a,color=a,fill=a,y=y)) +
  # geom_violin() +
  stat_summary(fun.y = median, geom = "point",shape=23,size=3) +
  stat_summary(fun.y = mean, geom = "point") +
  stat_summary(fun.data = mean_se, geom = "errorbar")

#NOTE: comparing using means and errorbars requires NORMAL DISTRIBUTION.

```

# Comparing Groups

The mean is a very popular number used to compare groups.

Unbiased Sample Variance
$$s_{N - 1} = \sqrt {\frac{1}{{N - 1}}\sum\limits_{i = 1}^N {\left( {x_i - \bar x} \right)^2 } }$$
