---
title: "The Tidyverse - A Universe of Coding Happiness and Joy"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}

library(tidyverse)

#OR load individual packages from the above metapackage
library(readxl)
library(dplyr)
library(tidyr)
library(stringr) #for dealing with strings
library(purrr) #for next time... replaces for/while loops
library(ggplot2)


```

# Why the tidyverse?

Base R functions are not always consistent and clear with what they expect as arguments and what they return. Packages in the `tidyverse` are designed to follow a consitent coding structure and work well together.

This means you only need to learn on *sub-language* for your entire data workflow!

## Import the data

```{r}

#look at sheets in spreadsheet
readxl::excel_sheets("data")

#load the sheet you want!
df <- readxl::read_excel("data",sheet = "specific")

#sweet nibblets!
df
```

## Manipulate the data!

* The most common useful functions are **filter**, **select** and **mutate**. Don't forget the pipe! **%>%** is the pipe. It is an *infix* function.

```{r}
  
manipulated <- df %>%
  select(something,another,mmmhmmm) %>% #select the columns you are interested in
  mutate( #create NEW variables/columns with mutate!
    rodfdw = str_extract(tower,"[A-D]"), #regular expression - used for text processing
    adf = str_extract(something,"[0-9].") %>% as.numeric(), #use stringr functions to manipulate strings!!!
    # grams_site = as.numeric(grams_site)
  ) %>%
  filter(adsf>0)

```

## Manipulate the data!

* ggplot works in **layers**. You add a layer with a `+`.

* Think about putting your response variable on the **y-axis** and looking at how it varies when categorized differently by using **x-axis**, **color** or maybe **size** or **facetting**.

* Graphing is a complicated business!!! There are TONS of ways to do things and change visuals. [stackoverflow](https://stackoverflow.com/) is you **best friend**. Also use the ? mark eg. `?ggplot`.

```{r}

manipulated %>%
  ggplot(aes(asdf,adf,fill=adsf)) +
  # geom_violin() +
  geom_boxplot() + 
  # ggpubr::stat_cor() +
  geom_blank()

```

## Where Next?

[R for Data Science](https://r4ds.had.co.nz/) is a phenomenal resource!!!!!!!!!!!!!!1